Community Analysis

### 2022-06-24 R/09-community-analysis.Rmd
rename communty analysis file

```{r initialize_packages}
library(tidyverse)
library(phyloseq)
library(decontam)
library(microViz)
library(knitr)
library(corncob)
library(DivNet)
library(magrittr)
library(vegan)
library(ggpubr)
# library(SpiecEasi)
library(ggthemes)
library(broom)

@@

```
There is no systematic difference in sequencing depth/library size between treatments. It does appear that sequencing depth decreases between passage 2 and 5.

@@

```{r, eval=FALSE}
richness_soil <- ps %>% breakaway
plot(richness_soil, physeq=ps, color="passage")
summary(richness_soil) %>% as_tibble

@@

save(bt_passage, file = 'Output/richness_model_passage.Rdata')

@@

save(bt_treat5, file = 'Output/richness_model_treat5.Rdata')

@@

save(bt_ch4, file = 'Output/richness_model_ch4.Rdata')

@@

Test confirms the above observation: Richness decreases between Passage 2
and Passage 5. No difference in richness by treatment in passage 5 or by
methane.

@@

ps_fixed <-
ps %>%
 tax_fix(
  min_length = 4,
  unknowns = c(""),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )

ps_fixed <- 
  ps_fixed %>% 
  ps_mutate(treat_bin = factor(ifelse(treat=='p', 1,0))) %>% 
  ps_mutate(passage_bin = factor(ifelse(passage==5, 1,0))) %>% 
  ps_mutate(treat_pas = interaction(treat_bin, passage_bin))

@@

```{r plot-beta, results=FALSE, fig.width = 6, fig.height = 2}
bray <-
  ord_calc_bray %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

jacc <- 
  ord_calc_jacc %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_fixed(3.9/48.3) +
  # coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

@@

```{r plot-aitch, fig.width = 4, fig.height = 4}

ord_calc_aitch %>% 
  ord_plot(color = "treat",
           auto_caption = FALSE) +
  theme_tufte() +
  geom_rangeframe() +
  coord_fixed(5.5/43.6, clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

```{r permanova-aitch, eval=FALSE}
dist_aitch <-
  ps %>% 
  dist_calc(dist = "aitchison")

dist_permanova(dist_aitch, c("passage", "treat"), "passage * treat")

passage_groups <- factor(c(rep(1,22), rep(2,24)), labels = c("2","5"))

dist_aitch %>% 
  dist_get %>% 
  vegan::betadisper(passage_groups) %>% 
  permutest()

treat_groups <- factor(c(rep(1,12), rep(2,10), rep(1,12), rep(2,12)), labels = c("n","p"))

dist_aitch %>% 
  dist_get %>% 
  vegan::betadisper(treat_groups) %>% 
  permutest()

ps %>%
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison") %>% 
  dist_permanova("treat")

ps %>%
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison") %>% 
  dist_permanova("treat")

adonis2(dist_get(dist_aitch) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'))

dist_aitch2 <-
  ps %>% 
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch2, "treat")
groups <- factor(c(rep(1,12), rep(2,10)), labels = c("N","P"))

dist_aitch2 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()
dist_aitch5 <-
  ps %>% 
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch5, "treat")
groups <- factor(c(rep(1,12), rep(2,12)), labels = c("N","P"))

There is a significant main effect and interaction of passage and treatment on
taxonomic dissimilarity of the soil microbiome. In addition, I reject the null
hypothesis of homogeneity of group dispersions for passage, but not for treatment.
Inspecting the NMDS plot, from this I conclude that passaging between Passage 2
and 5 both altered the soil microbiome taxonomic composition and reduced its 
variability. In addition, there is a difference in centroid between the two
treatments in both Passage 2 and Passage 5, but no difference in dispersion.

```{r beta-test}
# ps_prop <- transform_sample_counts(ps, function(otu) otu/sum(otu))
# 
# fit_bray <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "bray")
# 
# fit_bray
# 
# fit_jacc <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "jacc", binary = TRUE)
# 
# fit_jacc
ps_clr <- microbiome::transform(ps, "clr")

@@

fit_adonis <- rbind(tidy(fit_aitch), tidy(fit_bin_jacc))
fit_adonis$term <- c('Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total',
      'Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total')
save(fit_adonis, file = 'Output/beta_model.Rdata')

@@

###########3

# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
#         data = 
#           as(
#             sample_data(
#               ps_filter(
#                 ps_prop, 
#                 passage == 5)
#             ), 
#             'data.frame'), 
#         method = "bray")
# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
#         data = 
#           as(
#             sample_data(
#               ps_filter(
#                 ps_prop, 
#                 passage == 5)
#             ), 
#             'data.frame'), 
#         method = "jaccard",
#         binary = TRUE)

@@

@@

First, I looked at which taxa are unique to treatments N and P. Several families
present in P and absent in N. Most of them had low prevalance (present in 3 or
fewer samples) and low abundance (median < 5 reads across samples). There were two families unique
to the Positive treatment with relatively high prevalence. This included an ASV
that was a member of the Bacteroidia Class with no lower taxonomic designation. 
This ASV had a prevalence of 10/12 and a median abundance of ~2 reads. The other
prevalent family was a member of the Silvanigrellaceae, a newly described family places in
its own order. This family was present in
all 12 samples and was represented by three OTUs: one with no lower designation, one from the genus
Silvanigrella and one from the genus Candidatus Turbobacter. Silvanigrella was
isolated from a temperate fresh water lake (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijsem.0.001965).

@@

```{r, eval=FALSE}

g5<- 
  ps_fixed %>% 
  ps_filter(passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

g2<- 
  ps_fixed %>% 
  ps_filter(passage == 2) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

unique(g5$Family)[!(unique(g5$Family) %in% unique(g2$Family))]
unique(g5$Family)[unique(g5$Family) %in% unique(g2$Family)]

unique(g2$Family)[!(unique(g2$Family) %in% unique(g5$Family))]
unique(g2$Family)[unique(g2$Family) %in% unique(g5$Family)]

@@

#save(ps, file = 'ps.RData')
# Differential abundance test
corncob <- bbdml(formula = ASV2954 ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "lrt", boot = TRUE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05)
da_analysis <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "Wald", boot = FALSE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05, B = 50)
plot(da_analysis, level = c("Family"))
da_analysis$significant_models
dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)

sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
ntaxa(sig_taxa_treat)
save(da_analysis, sig_taxa_treat, file = 'Output/da_corncob.Rdata')

save(dt_data, file = 'Output/dt_data.Rdata')

@@

```{r stahp}
knitr::opts_chunk$set(eval = FALSE)

@@

```{r stacked_bar, eval=FALSE}
level = "Family"
ps_phylum <- speedyseq::tax_glom(ps, taxrank=level)
top10 = names(sort(taxa_sums(ps_phylum), TRUE)[1:10])
ps_phylum10 <- prune_taxa(top10, ps_phylum)
ps_phylum10 %>% 
  ps_filter(passage == 5) %>% 
  speedyseq::plot_bar(facet_grid = 'treat', fill = level) + 
  geom_bar(aes_string( color=level, fill=level), stat="identity", position="stack") +
  geom_bar(aes(x = reorder(rep, rank_ch4)))

## ALDex2

```{r, eval = FALSE}
library(ALDEx2)
data(selex)
#subset only the last 400 features for efficiency
selex.sub <- selex[1:400,]
conds <- c(rep("NS", 7), rep("S", 7))
x <- aldex.clr(selex.sub, conds, mc.samples=16, denom="iqlr", verbose=F)
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
head(x.effect)

```{r, eval=FALSE}
ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")

ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
ps_pass_5 <- microViz::ps_filter(ps_fixed, passage == 5)
otu_table <- as(t(otu_table(ps_fam_pas5_prev)), 'matrix')
conds <- c(rep("n", 12), rep("p", 12))

# Based on effect confidence interval

x <- aldex.clr(otu_table, conds, mc.samples=128, denom="all", verbose=F)
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

sgn <- sign(x.effect$effect.low) == sign(x.effect$effect.high)
par(mfrow=c(1,2))
plot(x.effect$rab.all, x.effect$diff.btw, pch=19, cex=0.3, col="grey", xlab="Abundance", ylab="Difference", main="Bland-Altman")
points(x.effect$rab.all[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
points(x.effect$rab.all[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")

plot(x.effect$diff.win, x.effect$diff.btw, pch=19, cex=0.3, col="grey",xlab="Dispersion", ylab="Difference", main="Effect")
points(x.effect$diff.win[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
points(x.effect$diff.win[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
abline(0,2, lty=2, col="grey")
abline(0,-2, lty=2, col="grey")

merge(as(tax_table(ps_fam_pas5_prev)[rownames(x.effect[sgn,]), ], 'data.frame'), x.effect[sgn,], by = 0)
```

```{r, eval=FALSE}

# Based on effect size

x.all <- aldex(otu_table, conds, mc.samples=128, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE)
effect_gt_2 <- merge(as(tax_table(ps_pass_5)[rownames(x.all[abs(x.all$effect)>2,]), ], 'data.frame'), x.all[abs(x.all$effect)>2,], by = 0)

effect_gt_2 %>% 
  ggplot(aes(x = effect, y = reorder(Family,effect))) +
  geom_point()

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

# identify which values are significant in both the t-test and glm tests
found.by.all <- which(x.all$we.eBH < 0.05 & x.all$wi.eBH < 0.05)

# identify which values are significant in fewer than all tests
found.by.one <- which(x.all$we.eBH < 0.05 | x.all$wi.eBH < 0.05)

# plot the within and between variation of the data
plot(x.all$diff.win, x.all$diff.btw, pch=19, cex=0.3, col=rgb(0,0,0,0.3),
 xlab="Dispersion", ylab="Difference")
points(x.all$diff.win[found.by.one], x.all$diff.btw[found.by.one], pch=19,
 cex=0.7, col=rgb(0,0,1,0.5))
points(x.all$diff.win[found.by.all], x.all$diff.btw[found.by.all], pch=19,
 cex=0.7, col=rgb(1,0,0,1))
abline(0,1,lty=2)
abline(0,-1,lty=2)

#View(tax_table(ps_fam_pas5_prev)[rownames(otu_table[found.by.all, ]), ])
```

```{r, eval=FALSE}
boxData <-
  ps_fixed %>% 
  ps_filter(passage ==5) %>% 
  tax_agg('Family') %>% 
  ps_get %>% 
  ps_melt()

Family_order<-
  boxData %>% 
  group_by(Family) %>% 
  summarize(median = median(Abundance)) %>% 
  arrange(median)
boxData$Family <- factor(boxData$Family, levels = Family_order$Family)
boxData$treat <- factor(boxData$treat, levels = c('p', 'n'))
```

```{r, eval=FALSE, fig.height=3, fig.width=12}

boxData %>% 
  ggplot(aes(x = Family, 
             y = Abundance)) +
  facet_grid(treat ~ .) +
  theme_tufte() +
  geom_rangeframe(color = "black") +
  # scale_x_reordered() +
  scale_y_continuous(trans = "pseudo_log") +
  geom_tufteboxplot() +
  theme(axis.text.x = element_text(angle = 90))
  # geom_boxplot() +
  # scale_color_discrete(guide = 'none') +
  # coord_flip()

```{r, eval=FALSE}
g5p<- 
  ps_fixed %>% 
  ps_filter(treat == "p", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()
g5n<- 
  ps_fixed %>% 
  ps_filter(treat == "n", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

g5p_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24))

g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12))

g5n_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 37))

g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 6))
```

## Faprotax on DA taxa

```{r export_to_faprotax, eval=FALSE}
faprotax_taxonomy <-
  sig_taxa_treat %>% 
  tax_table %>% 
  as_tibble %>% 
  mutate(.taxonomy = 
           faprotax_taxonomy %>% 
           select(Kingdom:Species) %>% 
           apply(1 , paste , collapse = "; " )) %>% 
  select(.otu, .taxonomy)

otu<-t(as(otu_table(sig_taxa_treat),"matrix")) 
  # 't' to transform if taxa_are_rows=FALSE
  #if taxa_are_rows=TRUE
  #otu<-as(otu_table(GlobalPatterns),"matrix"))
rownames(otu) <- faprotax_taxonomy$.taxonomy
otu_biom <- biomformat::make_biom(data=otu)
biomformat::write_biom(otu_biom,"otu_biom.biom")

```{bash faprotax, eval=FALSE}
conda activate faprotax

rm Output/func_table_from_biom.biom \
    Output/faprotax_report_from_biom.txt

python python/FAPROTAX_1.2.4/collapse_table.py \
    -i otu_biom.biom \
    -g python/FAPROTAX_1.2.4/FAPROTAX.txt \
    -o Output/func_table_from_biom.biom \
    -r Output/faprotax_report_from_biom.txt \
    -v

biom convert \
    -i Output/func_table_from_biom.biom \
    -o Output/func_table_from_biom_json.biom \
    --table-type="OTU table" \
    --to-json

conda deactivate
```

```{r, eval=FALSE}
biom_file <- "Output/func_table_from_biom_json.biom"

ps_faprotax <-
  biom_file %>% 
  biomformat::read_biom() %>% 
  biomformat::biom_data() %>% 
  as("matrix") %>% 
  otu_table(taxa_are_rows=TRUE) %>% 
  phyloseq(sample_data(ps_gen_pas5_prev))

da_faprotax <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ treat,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ treat,
                                 test = "Wald", boot = FALSE,
                                 data = ps_faprotax,
                                 fdr_cutoff = 0.05)
plot(da_faprotax)

## Picrust on DA taxa

```{r}
x = NULL
for (i in seq_along(da_analysis$significant_models)) {
  x <- c(x, da_analysis$significant_models[[i]]$coefficients[2, 1])
}
x > 0
sig_taxa_pos_neg <-
  da_analysis$significant_taxa %>% 
  as_tibble_col(column_name = ".otu") %>% 
  mutate(estimate = x) %>% 
  mutate(pos_or_neg = if_else(x > 0, "pos", "neg"))

top_da_taxa <- 
  sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>% 
  filter(.abundance > 2000) %>% 
  distinct(.otu) %$% .otu

taxonomy <-
  sig_taxa_treat %>% 
  tax_table %>% 
  as_tibble %>%
  filter(.otu %in% top_da_taxa)

taxonomy$.taxonomy <- taxonomy %>% 
           select(Kingdom:Species) %>% 
           apply(1 , paste , collapse = "; " )

taxonomy <-
  taxonomy %>% 
  select(.otu, .taxonomy)

df_top_da_taxa <-
  sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu != 'ASV4174') %>% 
  filter(.otu %in% top_da_taxa) %>% 
  left_join(taxonomy)

df_top_da_taxa %>% 
  ggplot(., aes(y = Genus, x = .abundance, color = treat)) +
  geom_jitter(height = 0.1) + 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ pos_or_neg)
ggsave('Output/dif_abund_results.png', width = 6, height = 4)

major_otus <- 
sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu != 'ASV4174') %>% 
  filter(.abundance > 300) %>% 
  select(.otu, pos_or_neg) %>% 
  distinct %>% 
  mutate(tax = otu_to_taxonomy(.otu, sig_taxa_treat))

otu_to_taxonomy(major_otus$.otu, sig_taxa_treat)

otu_to_taxonomy('ASV192', sig_taxa_treat)

sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu == 'ASV192') %>% 
  ggplot(., aes(x = .otu, y = .abundance, color = treat)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ pos_or_neg)

beijer_otu <-
  ps %>% 
  otu_table %>% 
  as_tibble %>% 
  filter(.otu == 'ASV192')

beijer_metadata <-
  ps %>% 
  sample_data %>% 
  as_tibble %>% 
    left_join(beijer_otu)

ggplot(beijer_metadata, aes(x = .abundance, y = log_ch4, color = treat)) +
  geom_point() +
  facet_wrap(~ passage)

df_top_da_taxa %>% 
filter(pos_or_neg == "pos") %>% 
select(Phylum:Genus) %>% 
distinct %>% 
  write_tsv(., 'top_positive_taxa.txt')
```

```{bash, eval=FALSE}
mkdir picrust2
```

```{r export_to_picrust}

top_da_taxa_pos <-
  df_top_da_taxa %>% 
  select(.otu, pos_or_neg) %>% 
  distinct(.otu, pos_or_neg) %>% 
  filter(pos_or_neg == "pos") %$%
  .otu
sig_taxa_treat %>% 
  tax_select(top_da_taxa_pos) %>% 
  otu_table %>% 
  as("matrix") %>% 
  t %>% 
  biomformat::make_biom() %>% 
  biomformat::write_biom("Output/study_seqs.biom")

sequences <-
  sig_taxa_treat %>% 
  tax_select(top_da_taxa_pos) %>% 
  refseq %>% 
  as.vector %>% 
  as.list
seqinr::write.fasta(sequences, names(sequences), file.out = "Output/study_seqs.fna")

```{bash picrust, eval=FALSE}
conda activate picrust2
cd picrust3c

place_seqs.py -s ../Output/study_seqs.fna \
    -o out.tre \
    -p 1 \
    --intermediate intermediate/place_seqs \
    --chunk_size 500
hsp.py -i 16S -t out.tre -o marker_predicted_and_nsti.tsv.gz -p 1 -n

hsp.py -i EC -t out.tre -o EC_predicted.tsv.gz -p 1

metagenome_pipeline.py -i ../Output/study_seqs.biom -m marker_predicted_and_nsti.tsv.gz -f EC_predicted.tsv.gz \
                       -o EC_metagenome_out --strat_out

pathway_pipeline.py -i EC_metagenome_out/pred_metagenome_contrib.tsv.gz \
                    -o pathways_out -p 1

add_descriptions.py -i EC_metagenome_out/pred_metagenome_unstrat.tsv.gz -m EC \
                    -o EC_metagenome_out/pred_metagenome_unstrat_descrip.tsv.gz

add_descriptions.py -i pathways_out/path_abun_unstrat.tsv.gz -m METACYC \
                    -o pathways_out/path_abun_unstrat_descrip.tsv.gz

#####

picrust2_pipeline.py -s Output/study_seqs.fna -i Output/study_seqs.biom -o picrust2_out_pipeline -p 4 --stratified

conda deactivate
```

```{r}
se.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=50))
se.gl.amgut <- spiec.easi(amgut1.filt, method='glasso', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=50))
sparcc.amgut <- sparcc(amgut1.filt)
## Define arbitrary threshold for SparCC correlation matrix for the graph
sparcc.graph <- abs(sparcc.amgut$Cor) >= 0.3
diag(sparcc.graph) <- 0
library(Matrix)
sparcc.graph <- Matrix(sparcc.graph, sparse=TRUE)
## Create igraph objects
ig.mb     <- adj2igraph(getRefit(se.mb.amgut))
ig.gl     <- adj2igraph(getRefit(se.gl.amgut))
ig.sparcc <- adj2igraph(sparcc.graph)

```{r}
library(igraph)
## set size of vertex proportional to clr-mean
vsize    <- rowMeans(clr(amgut1.filt, 1))+6
am.coord <- layout.fruchterman.reingold(ig.mb)

par(mfrow=c(1,3))
plot(ig.mb, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="MB")
plot(ig.gl, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="glasso")
plot(ig.sparcc, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="sparcc")

library(phyloseq)
## Load round 2 of American gut project
data('amgut2.filt.phy')
se.mb.amgut2 <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
                           nlambda=20, pulsar.params=list(rep.num=50))
ig2.mb <- adj2igraph(getRefit(se.mb.amgut2),  vertex.attr=list(name=taxa_names(amgut2.filt.phy)))
plot_network(ig2.mb, amgut2.filt.phy, type='taxa', color="Rank3")

ps_p <-
  test %>% 
  speedyseq::tax_glom("Family") %>% 
  ps_filter(passage == "5") %>% 
  ps_filter(treat == "p")

se.mb.ps_p <- spiec.easi(ps_p, method='mb', lambda.min.ratio=1e-2,
                           nlambda=200, pulsar.params=list(rep.num=50))
ig2.mb.ps_p <- adj2igraph(getRefit(se.mb.ps_p),  vertex.attr=list(name=taxa_names(ps_p)))
plot_network(ig2.mb.ps_p, ps_p, type='taxa', color = "Class")

ps_n <-
  test %>% 
  speedyseq::tax_glom("Family") %>% 
  ps_filter(passage == "5") %>% 
  ps_filter(treat == "n")

se.mb.ps_n <- spiec.easi(ps_n, method='mb', lambda.min.ratio=1e-2,
                           nlambda=200, pulsar.params=list(rep.num=50))
ig2.mb.ps_n <- adj2igraph(getRefit(se.mb.ps_n),  vertex.attr=list(name=taxa_names(ps_n)))
plot_network(ig2.mb.ps_n, ps_n, type='taxa', color="Class")
getOptMerge(se.mb.ps_p)
getOptMerge(se.mb.ps_n)
getStability(se.mb.ps_n)
###

se1.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
se2.mb.amgut <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))

se1.slr.amgut <- spiec.easi(amgut1.filt, method='slr', r=10, lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
se2.slr.amgut <- spiec.easi(amgut2.filt.phy, method='slr', r=10, lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))

>>>>

```{r initialize_packages}
library(tidyverse)
library(phyloseq)
library(decontam)
library(microViz)
library(knitr)
library(corncob)
library(DivNet)
library(magrittr)
library(vegan)
library(ggpubr)
library(ggthemes)
library(broom)

@@

There is no systematic difference in sequencing depth/library size between treatments. It does appear that sequencing depth decreases between passage 2 and 5.

@@

```{r richness}

# Test effect of passage number on richness using all samples
richness_soil <- ps %>% breakaway
plot(richness_soil, physeq=ps, color="passage")
summary(richness_soil) %>% as_tibble

@@

# Test effect of treatment on richness in passage 5

@@

# Test effect of CH4 on richness

@@

save(bt_passage, bt_treat5, bt_ch4, file = 'Output/richness_models.RData')

@@

Test confirms the above observation: Richness decreases between Passage 2
and Passage 5. No difference in richness by treatment in passage 5 or by
methane flux.

@@

# Replace unknown and missing taxa
ps_fixed <-
ps %>%
 tax_fix(
  min_length = 4,
  unknowns = c(""),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )

# Recode treatments as binary 1/0
ps_fixed <- 
  ps_fixed %>% 
  ps_mutate(treat_bin = factor(ifelse(treat=='p', 1,0))) %>% 
  ps_mutate(passage_bin = factor(ifelse(passage==5, 1,0))) %>% 
  ps_mutate(treat_pas = interaction(treat_bin, passage_bin))

@@

## Bray-Curtis and Jaccard

```{r plot-beta, results=FALSE, fig.width = 8, fig.height = 3}
bray <-
  ord_calc_bray %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2')) +
  labs(title = "Bray-Curtis")

jacc <- 
  ord_calc_jacc %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_fixed(3.9/48.3) +
  # coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2')) +
  labs(title = "Jaccard")

@@

## Aitchison

```{r beta-test}

ps_clr <- microbiome::transform(ps, "clr")

@@

fit_adonis <- rbind(tidy(fit_aitch), tidy(fit_bin_jacc))
fit_adonis$term <- c('Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total',
      'Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total')
save(fit_adonis, file = 'Output/beta_model.RData')

@@

@@

There is a significant main effect and interaction of passage and treatment on
taxonomic dissimilarity of the soil microbiome. In addition, there is no difference
in dispersion between treatments, but there is between passage2 and 5.

@@

First, I looked at which taxa are unique to treatments N and P. Several families
present in P and absent in N. Most of them had low prevalance (present in 3 or
fewer samples) and low abundance (median < 5 reads across samples). There were two families unique
to the Positive treatment with relatively high prevalence. This included an ASV
that was a member of the Bacteroidia Class with no lower taxonomic designation. 
This ASV had a prevalence of 10/12 and a median abundance of ~2 reads. The other
prevalent family was a member of the Silvanigrellaceae, a newly described family placed in
its own order. This family was present in
all 12 samples and was represented by three OTUs: one with no lower designation, one from the genus
Silvanigrella and one from the genus Candidatus Turbobacter. Silvanigrella was
isolated from a temperate fresh water lake (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijsem.0.001965).

@@

```{r}
# families in to Passage 5, Treatment P
g5p<- 
  ps_fixed %>% 
  ps_filter(treat == "p", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

# families in to Passage 5, Treatment N
g5n<- 
  ps_fixed %>% 
  ps_filter(treat == "n", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

# Plot families unique to P5 in order of median abundance
g5p_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24)) +
  labs(title = "Unique to P5 by abundance")

# Plot families unique to P5 in order of prevalence
g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12)) +
  labs(title = "Unique to P5 by prevalence")

# Plot families unique to N5 in order of abundance
g5n_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 37)) +
  labs(title = "Unique to N5 by abundance")

# Plot families unique to N5 in order of prevalence

g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 6)) +
  labs(title = "Unique to N5 by prevalence")

@@

# Differential abundance test
da_analysis <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "Wald", boot = FALSE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05, B = 50)

#plot(da_analysis, level = c("Family"))
da_analysis$significant_models
dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)

sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
ntaxa(sig_taxa_treat)
save(da_analysis, sig_taxa_treat, dt_data, file = 'Output/da_corncob.RData')

@@

```{r stop}
knitr::opts_chunk$set(eval = FALSE)

@@
